Pre-trained encoders for dialogue: The case of laughter

### 2019-12-07 reports/ACL2020 Paper/acl2020-paper.tex
Fixes according to Staffan's comments

\begin{document}
\maketitle
\begin{abstract}
  The degree to which language models pre-trained on text data are useful for dialogue is an open question.
  In this paper, we investigate the usefulness of BERT for dialogue act recognition.
  We are interested in both how useful the pre-training proceedure is and 
  how well the model adapts to the dialogue domain.
  To confirm that the fine-tuned model learns to represent dialogical features, 
  we look at how it uses laughter, a phenomenon specific to dialogue.
\end{abstract}

@@

Dialogue, especially spoken dialogue, is radically different from the kind of data that neural language models are pre-trained on.
In the example from the Switchboard corpus, shown in table~\ref{table:example}, it is evident that the syntactic structure of dialogue is very different from  written text, even if one imagines a script for a movie or a play.
Not only is the internal structure of contributions different---with features such as disfluencies, repair, incomplete sentences, and various vocal sounds---but the sequential structure of the discourse is also very different.
In dialogue, speakers take turns and switch perspectives.
More generally, the way that utterances contribute and cohere to the discourse is different from the relationship between sentences in a piece of written text.

One important dialogical feature missing from text data is laughter.
In the Switchboard dialogue corpus, laughter comes about every 200 tokens.
Laughter also relates to the discourse structure of dialogue and can relate to a laughable, which can be any an event or an entity in the discourse (and external perceived event as well).
Laughter can precede, follow or overlap the laughable, with the time alignment between the laughter and laughable dependent on who produces the laughable and the form of the laughter \citep{tian2016we}.
In this work we found out that laughs can also be helpful to disambiguate a dialogue act.

@@

The distribution of laughs in different dialogue acts has a rather uniform shape with a few outliers (see box plots). Figure \ref{fig:by-da} in Appendix~\ref{sec:suppl} provides a breakdown of dialogue act depending on whether current or adjacent utterances (the preceding utterance and the following one are counted as adjacent) contain laughter.\footnote{Note that SWDA has a ``Nonverbal'' dialogue act which is misguiding with respect to laughter, because utterances only containing a single laughter token fall into this category, however they can serve, for example, to acknowledge a statement or to deflect a question  \citep{mazzocconi2019phd}.}

\begin{figure}
  \centering
  \includegraphics[width=0.8\linewidth]{img/box-swda.pdf}
  \caption{Box plot for utterances associated with laughter in SWDA. On the left: proportion of DAs having laughter in one of the adjacent utterances, on the right: proportion of DAs containing laughter. }
    \label{fig:box-swda}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[width=0.8\linewidth]{img/box-ami.pdf}
  \caption{Box plot for utterances associated with laughter in AMI-DA. On the left: proportion of DAs having laughter in one of the adjacent utterances, on the right: proportion of DAs containing laughter. }
    \label{fig:box-ami}
\end{figure}

For SWDA corpus we also considered grouping the dialogue acts according to DAMSL coders manual \citep{jurafskySwitchboardSWBDDAMSLShallowDiscourseFunction1997a}: Forward-Communicative-Function (51.86\% of all DAs), Backwards-Communicative-Function (30.35\%), Other (9.04\%) and Communicative-Status (8.75\%). We observed even distribution of laughs in these groups, except the Communicative-Status group which contains ``Nonverbal'' dialogue acts.

@@

Confusion matrices (Figure \ref{fig:swda-cm}) provide some food for thought.
One of the most obvious things that can be observed is rhetorical questions: supposedly laughter, can be used as a device to cancel seriousness of the question.
Therefore it can help to disambiguate such questions like the one we show in example in Table~\ref{table:example-qh}.
\begin{table}
      \small
  \centering
  \begin{tabularx}{\linewidth}{llL}
    \toprule
    Speaker & DA & Utterance \\ \midrule
        \multicolumn{3}{c}{\emph{(Talking about hobbies)}}\\
    B   & sd	&  Um, as far as spare time, they talked about, \\
    B	& \% & I don't, + I think, \\
    B	& qh & who has any spare time \texttt{<laughter>}? \\
    A	& x & \texttt{<laughter>}.\\
             \bottomrule
  \end{tabularx}
  \caption{Example from the SWDA corpus (sw3735). B's contribution \emph{qh}---Rhetorical question is misinterpreted as \emph{qw}---Wh-question by the BERT model without laughs in training data. }
  \label{table:example-qh}
\end{table}

  \begin{figure}
  \centering
  \includegraphics[width=0.9\linewidth]{img/swda-cm.pdf}
  \caption{Confusion matrices for \texttt{BERT-NL} (top) vs \texttt{BERT-L} (bottom); SWDA corpus.}
    \label{fig:swda-cm}
  \end{figure}

In order to further investigate the differences between the models and datasets, we looked at the dialogue act groups.
Mapping all dialogue acts into four groups allows direct comparison between model performances in SWDA and AMI-DA.
Table~\ref{table:laughter-group-acc} compares accuracies between the two main groups: Forward-Communicative-Function and Backward-Communicative-Function (82.3\% of all DAs in SWDA and 76.7\% in AMI-DA).
Our models tend to perform better for Forward-Communicative-Function group irrespective of presence the corpus and the presence of laughs.
Also, our models perform better on the AMI-DA corpus.
Overall performance in these groups appears to be unaffected by presence of laughs.

@@

\section{Discussion} % add items to discuss
\begin{itemize}
  \item RoBERTa training scheme finds that next sentence prediction is of little importance. Is this true for us too? (suspect not due to discourse importance). see: https://github.com/huggingface/transformers/issues/1622
  \item Issues related to working with transcribed speech: Different kinds of laughter are all represented by the same token. Does transcription orthography capture prosody? We know prosidy is useful for DAR (Jurafsky).
  \item future work: pre-training on very large un-labeled dialogue corpora. How does text chat transfer to the spoken/transcribed setting?
\end{itemize}

\bibliography{acl2020}
\bibliographystyle{acl_natbib}
\clearpage
\appendix
% to appendix?
\begin{figure*}[h!]
  \section{Supplementary materials}\label{sec:suppl}
  \centering
  \includegraphics[width=\linewidth]{img/SWDA-bertLvsNL.pdf}
  \includegraphics[width=\linewidth]{img/AMI-DA-bertLvsNL.pdf}
  \caption{Change in accuracy for each dialogue act. Positive changes when adding laughter are shown in blue. Vertical bars indicate how often dialogue act is associated with laughter. Top chart: SWDA, Bottom chart: AMI-DA. }
    \label{fig:by-da}
  \end{figure*}

>>>>

\begin{document}
\maketitle
\begin{abstract}
  The degree to which language models pre-trained on text data are useful for dialogue is an open question.
  In this paper, we investigate the usefulness of BERT for dialogue act recognition.
  We are interested in both how useful the pre-training procedure is and 
  how well the model adapts to the dialogue domain.
  To confirm that the fine-tuned model learns to represent dialogical features, 
  we look at how it uses laughter, a phenomenon specific to dialogue.
\end{abstract}

@@

Dialogue, especially spoken dialogue, is radically different from the kind of data that neural language models are pre-trained on.
In the example from the Switchboard corpus, shown in Table~\ref{table:example}, it is evident that the syntactic structure of dialogue is very different from  written text, even if one imagines a script for a movie or a play.
Not only is the internal structure of contributions different---with features such as disfluencies, repair, incomplete sentences, and various vocal sounds---but the sequential structure of the discourse is also very different.
In dialogue, speakers take turns and switch perspectives.
More generally, the way that utterances contribute to and cohere with the discourse is different from the relationship between sentences in a piece of written text.

One important dialogical feature missing from text data is laughter, which is ubiquitous in our everyday interactions.
We observe that in the Switchboard dialogue corpus laughter comes about every 200 tokens.
Laughter also relates to the discourse structure of dialogue and can relate to a laughable, which can be any an event or an entity in the discourse (and external perceived event as well).
Laughter can precede, follow or overlap the laughable, with the time alignment between the laughter and laughable dependent on who produces the laughable and the form of the laughter \citep{tian2016we}.
In this work we found out that laughs can also be helpful to disambiguate a dialogue act.

@@

\paragraph{Laughs in the data}
The distribution of laughs in different dialogue acts has a rather uniform shape with a few outliers (see box plots). Figure \ref{fig:by-da} in Appendix~\ref{sec:suppl} provides a breakdown of dialogue act depending on whether current or adjacent utterances (the preceding utterance and the following one are counted as adjacent) contain laughter.\footnote{Note that SWDA has a ``Nonverbal'' dialogue act which is misguiding with respect to laughter, because utterances only containing a single laughter token fall into this category, however they can serve, for example, to acknowledge a statement or to deflect a question  \citep{mazzocconi2019phd}.}

\begin{figure}
  \centering
  \includegraphics[width=0.9\linewidth]{img/box-swda.pdf}
  \caption{Box plot for dialogue acts associated with laughter in SWDA. On the left: proportion of DAs having laughter in one of the adjacent utterances, on the right: proportion of DAs containing laughter. }
    \label{fig:box-swda}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[width=0.9\linewidth]{img/box-ami.pdf}
  \caption{Box plot for dialogue acts associated with laughter in AMI-DA. On the left: proportion of DAs having laughter in one of the adjacent utterances, on the right: proportion of DAs containing laughter. }
    \label{fig:box-ami}
\end{figure}

\paragraph{Dialogue act grouping}
For our corpora we also considered grouping the dialogue acts according to DAMSL coders manual \citep{jurafskySwitchboardSWBDDAMSLShallowDiscourseFunction1997a}: Forward-Communicative-Function (51.86\% of all DAs), Backwards-Communicative-Function (30.35\%), Other (9.04\%) and Communicative-Status (8.75\%).
This allows the direct comparison between SWDA and AMI-DA.
With respect to laughs, we observed even distribution of them in the groups, except the Communicative-Status group in the case of SWDA, because it contains ``Nonverbal'' dialogue acts.

@@

Confusion matrices (Figure \ref{fig:swda-cm}) provide some food for thought.
% TODO: explain why most of misclassifications are sd (majority class)
One of the most obvious things that can be observed is rhetorical questions: supposedly, laughter can be used as a device to cancel seriousness \citep{ginzburg2015understanding} of the question.
Therefore it can help to disambiguate such questions like the one we show in example in Table~\ref{table:example-qh}.
\begin{table}
      \small
  \centering
  \begin{tabularx}{\linewidth}{llL}
    \toprule
    Speaker & DA & Utterance \\ \midrule
        \multicolumn{3}{c}{\emph{(Talking about hobbies)}}\\
    B   & sd	&  Um, as far as spare time, they talked about, \\
    B	& \% & I don't, + I think, \\
    B	& qh & who has any spare time \texttt{<laughter>}? \\
    A	& x & \texttt{<laughter>}.\\
             \bottomrule
  \end{tabularx}
  \caption{Example from the SWDA corpus (sw3735). B's contribution \emph{qh} (Rhetorical question) is misinterpreted as \emph{qw} (Wh-question) by the BERT model without laughs in training data. }
  \label{table:example-qh}
\end{table}

  \begin{figure}
  \centering
  \includegraphics[width=0.9\linewidth]{img/swda-cm.pdf}
  \caption{Confusion matrices for \texttt{BERT-NL} (top) vs \texttt{BERT-L} (bottom); SWDA corpus. Solid line shows the improvement in classification of rhetorical questions.}
    \label{fig:swda-cm}
  \end{figure}

In order to further investigate the differences between the models and datasets, we looked at the dialogue act groups.
Mapping all dialogue acts into four groups allows direct comparison between model performances in SWDA and AMI-DA.
Table~\ref{table:laughter-group-acc} compares accuracies between the two main groups: Forward-Communicative-Function and Backward-Communicative-Function (82.3\% of all DAs in SWDA and 76.7\% in AMI-DA).
Our models tend to perform better for Forward-Communicative-Function group for both corpora irrespective of the presence of laughs.
Also, our models perform better on the SWDA corpus.
Overall performance in these groups appears to be unaffected by presence of laughs.

@@

\section{Discussion} % add items to discuss
\begin{itemize}
  \item RoBERTa training scheme finds that next sentence prediction is of little importance. Is this true for us too? (suspect not due to discourse importance). see: https://github.com/huggingface/transformers/issues/1622
  \item Issues related to working with transcribed speech: Different kinds of laughter are all represented by the same token. Does transcription orthography capture prosody? We know prosidy is useful for DAR (Jurafsky).
  \item future work: pre-training on very large un-labeled dialogue corpora. How does text chat transfer to the spoken/transcribed setting?
  \item Laughter is more helpful in SWDA than in AMI-DA. This observation may support the hypothesis that in SWDA due to the nature of interactions over the phone, dialogue participants can't rely on visual signals, such as gestures and facial expressions, therefore vocalisations, such as laughter, are more pronounced and therefore more helpful to disambiguate the utterances.
  \item The point above can also explain why our models better on SWDA: just a larger proportion of stuff that interlocutors rely on is in the transcription.
\end{itemize}

\bibliography{acl2020}
\bibliographystyle{acl_natbib}
\clearpage
\appendix
\begin{figure*}[h!]
  \section{Supplementary materials}\label{sec:suppl}
  \centering
  \includegraphics[width=\linewidth]{img/SWDA-bertLvsNL.pdf}
  \includegraphics[width=\linewidth]{img/AMI-DA-bertLvsNL.pdf}
  \caption{Change in accuracy for each dialogue act. Positive changes when adding laughter are shown in blue. Vertical bars indicate how often dialogue act is associated with laughter. Top chart: SWDA, Bottom chart: AMI-DA. }
    \label{fig:by-da}
  \end{figure*}
